---
title: "Prediction Assignment"
author: "Tom -vdM"
date: "15 november 2019"
output: html_document
---
##Introduction
The goal of is to predict the manner in which the subjects did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. 


###Setting seed, loading packages and data
```{r}
rm(list=ls())
set.seed(1234)

require(caret)
require(randomForest)
require(rpart)
require(rpart.plot)

trainLink <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testLink <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(trainLink, na.strings = c("NA", "#DIV/0!", ""))
testing <- read.csv(testLink, na.strings = c("NA", "#DIV/0!", ""))
```

###Cleaning data and overview
```{r}
training <- training[,colSums(is.na(training)) == 0]
testing <- testing[,colSums(is.na(testing)) == 0]

training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]

head(training)
dim(training) #19662 cases, 53 variables
```
###Subsetting training data for cross-validation
```{r}
split <- createDataPartition(y=training$classe, p=0.75, list=FALSE)

trainTrain <- training[split,]
trainTest <- training[-split,]

nrow(trainTrain)
nrow(trainTest)

summary(trainTrain$classe) #similar distribution over levels, with peak in "A"
```
###Predicting1; Decision Tree
```{r}
mod1 <- rpart(classe ~ ., data = trainTrain, method = "class") #we use all variables with relevant (non-NA) information
pred1 <- predict(mod1, trainTest, type = "class")
rpart.plot(mod1) #overview; hard to read due to number of variables included in analyses

confusionMatrix(pred1, trainTest$classe)
```
Outcome: The decision tree analysis has a accuracy of 0.7394 [0.73; 0.75]

###Predicting2; Random Forest
```{r}
mod2 <- randomForest(classe ~ ., data = trainTrain, method = "class")
pred2 <- predict(mod2, trainTest, type = "class")

confusionMatrix(pred2, trainTest$classe)
```
Outcome: The Random Forest has an accuracy of 0.9955 [0.99; 1.00]

###Model picking
Random Forest performed with a much higher accuracy --> method of choice

###Prediction of manner of exercise on test cases, based on Random Forest
```{r}
predTest <- predict(mod2, testing, type = "class")
predTest
```


